---
title: "Group_1"
author: "Joseph Lynch and Anand Taxali"
date: "5/17/2020"
output:
  html_document: default
  pdf_document: default
---



## Problem 1 Forest fires data

```{r fig.align = "center"}

library(ggplot2)
forestfires <- read.csv("forestfires.csv")
attach(forestfires)
str(forestfires)
monthly <- factor(forestfires$month, levels = c("jan", "feb", "mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec"))
daily <- factor(forestfires$day, levels = c("mon", "tue", "wed", "thu", "fri", "sat", "sun"))

# 1a Plot area vs.temp, area vs. month, area vs. DC, area vs. RH for January
# through December combined in one graph. Hint: Place area on Y axis and use
# 2x2 matrix to place the plots adjacent to each other.

par(mfcol = c(2, 2))
plot(temp, area,  xlab = "Temperature", ylab = "Area")
plot(monthly, area,  xlab = "Month", ylab = "Area")
plot(DC, area,  xlab = "DC Index", ylab = "Area")
plot(RH, area,  xlab = "Relative Humidity", ylab = "Area")

```{r fig.align = "center"}
#1b b. Plot the histogram of wind speed (km/h).

hist(wind, xlab = "Wind speed(km/h)")

# 1c Compute the summary statistics (min, 1Q, mean, median, 3Q, max,) of part b.
summary(wind)

# 1d Add a density line to the histogram in part b.
hist(wind, xlab = "Wind speed(km/h)",freq=FALSE)
lines(density(wind), col="red", lwd=2)
```{r fig.align = "center"}
#1e Plot the wind speed density function of all months in one plot. Use
#different colors for different months in the graph to interpret your result clearly.
#[Hint: use ggplot + geom_density or qplot(geom=density)]
ggplot(data=forestfires, aes(x=wind,color=monthly)) +
  geom_density(alpha=.3)


#1f. Plot the scatter matrix for temp, RH, DC and DMC. How would you interpret the
#result in terms of correlation among these data?
library(gclus)
mydata <- forestfires[c("temp", "RH", "DC", "DMC")]
mydata.corr <- abs(cor(mydata))
myorder <- order.single(mydata.corr)
mycolors <- dmat.color(mydata.corr)
cpairs(mydata,
       myorder,
       panel.colors=mycolors,
       gap=.5,
       main="Variables Ordered and Colored by Correlation"
)
```{r fig.align = "center"}
#1f answers
# DMC has a strong positive correlation to DC, but no strong correlation to the other variables
# Temperature and RH have a strong correlation
# DC index and tempature have a strong correlation, with high DC=high temp and low DC=low temp


#1g Create boxplot for wind, ISI and DC. Are there any anomalies/outliers? Interpret
# your result. 
par(mfrow=c(1,3))
boxplot(wind,main="wind")
boxplot(ISI,main="ISI")
boxplot(DC,main="DC")

#1g answers 
# Both wind and ISI have some outliers outside of the ranges of the boxplot
# DC has a higher skewed distributio, with the median closer to Q3
```{r fig.align = "center"}

#1h. Create the histogram of DMC. Create the histogram of log of DMC. Compare the
#result and explain your answer
par(mfrow=c(1,2))
hist(DMC, xlab="DMC")
hist(log(DMC), xlab="log DMC")

#The distribution is right skewed with normal DMC and left skewed with log of DMC

```{r fig.align = "center"}

####################################################################################
#Problem 2 


data1 <- read.csv("M01_quasi_twitter.csv")
attach(data1)
#2a How are the data distributed for friend_count variable?

plot(density(data1$friends_count))
hist(data1$friends_count)
boxplot(data1$friends_count)

#The data is constricted in a small range having large values

#2b Compute the summery statistics (min, 1Q, mean, median, 3Q, max) on
#friend_count?

summary(data1$friends_count)
```{r fig.align = "center"}
#2c How is the data quality in friend_count variable?
#
#The data quality is not that great as the range is too big. also, it is not a normal
#distribution as the mean is bigger than the median.

#2d. Produce a 3D scatter plot with highlighting to impression the depth for variables
#below on M01_quasi_twitter.csv dataset. created_at_year, education, age.
#Put the name of the scatter plot "3D scatter plot".
library(plotrix)
scatterplot3d::scatterplot3d(data1$created_at_year, data1$education, data1$age,
                             xlab = "created_at_year",
                             ylab = "education",
                             zlab = "age",
                             highlight.3d=TRUE,
                             col.axis="black",
                             col.grid="yellow",
                             main="3D scatter plot",
                             pch=20,
                             cex.symbols=1,angle= 8)
```{r fig.align = "center"}
#2e. Consider 650, 1000, 900, 300 and 14900 tweeter accounts are in UK, Canada,
#India, Australia and US, respectively. Plot the percentage Pie chart includes 
#percentage amount and country name adjacent to it, and also plot 3D pie chart
#for those countries along with the percentage pie chart. Hint: Use C=(1, 2) matrix
#form to plot the charts together.
par(mfcol = c(1, 2))
countries <- c(14900,1000,650,300,900)

lbls <- c("USA", "Canada","United Kingdom","Australia","India")
pct <- round(countries/sum(countries)*100,1)
lblsUpdated <- paste(lbls, " ", pct, "%", sep = "") 
pie(countries,labels = lblsUpdated,main="Pie Chart of data1 Accounts")

pie3D(countries,labels=lblsUpdated,
      radius=1.0,
      labelcex=0.8,
      explode=0.,
      theta=1.0,
      main="Pie Chart of data1 Accounts")

#2f Create kernel density plot of created_at_year variable and interpret the result?

par(mfrow=c(1,1))
YearCreatedDensity <- density(data1$created_at_year)
plot(YearCreatedDensity, main="Kernel Density of Created at Year")
polygon(YearCreatedDensity, col="green", border="orange")
rug(data1$created_at_year, col="blue") 


#Based on the density plot, we can see that the values are mainly
#concentrated between 2008 and 2014 and where the probability of
#getting 2009 value being the highest.

```{r fig.align = "center"}



#####################################################################################



# Problem 3

#3a.Standardize the data and create new dataset with standardized data and name it
# Ndata.


InsuranceClaims <- read.csv("raw_data.csv")
attach(InsuranceClaims)



Ndata <- scale(InsuranceClaims,center=TRUE, scale=TRUE)
Ndata <- as.data.frame(Ndata)


```{r fig.align = "center"}
#3b. Create the boxplot of all the variables in their original form.

boxplot(InsuranceClaims,main="Original Form")


#3c. Create boxplot of all the variables in their standardized form

boxplot(Ndata,main="Standardized Form")

#3d
par(mfrow=(c(1,2)))
boxplot(InsuranceClaims,main="Original Form")
boxplot(Ndata,main="Standardized Form")

# In it's original form, we can see the data's distributions individually but they are more difficult to compare distribution-wise since they have different values.
# In standardized form, it is easier to compare the distributions of the four different values since they are on the same normal distribution scale. 

```{r fig.align = "center"}
#3e

attach(Ndata)
library(car)

scatterplot(Ndata$A~Ndata$B, xlab="Sustainability",ylab="Carbon Footprint")

#Sustainability and Carbon footprint to do not appear to have any correlation in this graph. 